DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-20 paper/overview.tex
Paper revisions

Note that although the steering angle output of the network ($angle$) is
a continuous real value, the RC car platform we used only supports
three discrete angles---left (-30$^{\circ}$), center 
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
We approximate the network generated real-valued angle to the closest
one of the three angles. Although this may introduce inaccuracy in
control, the observed control performance of the system is respectable,
likely due to the inherently stocastic nature of DNN.

@@

Another factors that can potentially affect the prediction accuracy of
the CNN, are camera and actuator (motor) control latencies. The camera
latency is from the time the camera sensor observes the scene to the
time the computer actually reads the digitized image data. This time
can be noticable depending on the camera used and the data processing
time of the computing platform. Higher camera latency could
negatively affect control performance, because the DNN would analyze
stale scenes. The actuator (motor) control latency from the time
the control output is sent to the steering motor to the time the motor
actually moved at a desired position, which also can takes
considerable time. In our platform, the combined latency is measured
to be around 50 ms, which is reasonble.
%% In other words, %% in the recorded video and control data, 
%% a control action of the CNN appears to be applied 50 ms later.
%% Considering that camera's framerate is 30Hz
%% (33.3 ms/frame), this is about two frames after the control action.
%% We experimentally measured the camera
%% latency and found it to be around 50-100 ms.
If this value is too high, control performance may suffer.
Our initial prototype suffered this problem as the observed latency
was as high as 300 ms, which negatively affected control performance.
For reference, the latency of human perception is known to be as fast
as 13 ms~\cite{ThomasBurger2015}. 
% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/

Our trained DNN models showed good prediction accuracy, successfully
navigating several different tracks we used for training.
For instance, the DeepPicar could remain on Track 1
(Figure \label{fig:track}) over 10 minutes at a moderate speed (50\%
throttle), at which point we stopped the experiment, and more than one
minute at a higher speed (75\% throttle)~\footnote{Self-driving videos: \url{https://photos.app.goo.gl/q40QFieD5iI9yXU42}
  %% \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}
}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).  Also, we
find that the prediction accuracy is significantly affected by the
quality of training data as well as various environmental factors such
as lighting conditions. We plan to investigate more systematic ways
to improve the CNN's prediction accuracies.

>>>>

Note that although the steering angle output of the network ($angle$) is
a continuous real value, the RC car platform we used only supports
three discrete angles---left (-30$^{\circ}$), center 
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
We approximate the network generated real-valued angle to the closest
one of the three angles. Although this may introduce inaccuracy in
control, the observed control performance of the system is respectable,
likely due to the inherently stocastic nature of CNNs.

@@

Another factors that can potentially affect the prediction accuracy of
the CNN, are camera and actuator (motor) control latencies. The camera
latency is from the time the camera sensor observes the scene to the
time the computer actually reads the digitized image data. This time
can be noticable depending on the camera used and the data processing
time of the computing platform. Higher camera latency could
negatively affect control performance, because the CNN would analyze
stale scenes. The actuator (motor) control latency from the time
the control output is sent to the steering motor to the time the motor
actually moved at a desired position, which also can takes
considerable time. In our platform, the combined latency is measured
to be around 50 ms, which is reasonble.
%% In other words, %% in the recorded video and control data, 
%% a control action of the CNN appears to be applied 50 ms later.
%% Considering that camera's framerate is 30Hz
%% (33.3 ms/frame), this is about two frames after the control action.
%% We experimentally measured the camera
%% latency and found it to be around 50-100 ms.
If this value is too high, control performance may suffer.
Our initial prototype suffered this problem as the observed latency
was as high as 300 ms, which negatively affected control performance.
For reference, the latency of human perception is known to be as fast
as 13 ms~\cite{ThomasBurger2015}. 
% https://www.pubnub.com/blog/2015-02-09-how-fast-is-realtime-human-perception-and-technology/

Our trained CNN models showed good prediction accuracy, successfully
navigating several different tracks we used for training.
For instance, the DeepPicar could remain on Track 1
(Figure \label{fig:track}) over 10 minutes at a moderate speed (50\%
throttle), at which point we stopped the experiment, and more than one
minute at a higher speed (75\% throttle)~\footnote{Self-driving videos: \url{https://photos.app.goo.gl/q40QFieD5iI9yXU42}
  %% \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}
}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).  Also, we
find that the prediction accuracy is significantly affected by the
quality of training data as well as various environmental factors such
as lighting conditions. We plan to investigate more systematic ways
to improve the CNN's prediction accuracies.

### 2018-04-20 paper/related.tex
Paper revisions

There are several relatively inexpensive RC-car based autonomous car
testbeds. MIT's RaceCar~\cite{shin2017project} and UPenn's
F1$/$10~\cite{upennf1tenth} are both based on a Traxxas 1/10 scale RC
car and a NVIDIA Jetson multicore computing platform, which
is equipped with many sophisticated sensor packages such as a lidar.
However, they both cost more than \$3,000, requiring considerable
investment. DonkeyCar~\cite{donkeycar} is similar to our DeepPicar as
it also uses a Raspberry Pi 3 and a similar CNN for end-to-end
control, although it costs more (about \$200).
%% because it uses a more
%% expensive 1/16 scale RC car than our 1/24 scale one.
The main contribution of our paper is in the detailed analysis of
computational aspects of executing a CNN-based real-time control
workload on diverse embedded computing platforms.

In this paper, we have analyzed real-time performance of a real-world
CNN, which was used in NVIDIA's DAVE-2 self-driving
car~\cite{Bojarski2016}, on a low-cost raspberry pi 3 quad-core 
platform and other embedded multicore platforms. It should be noted, 
however, that DAVE-2's CNN is relatively small compared
to recent state-of-the-art CNNs, which are increasingly larger and
deeper. For example, the CNN based object detector models evaluated in
Google's recent study~\cite{huang2017speed} have between 3M to 54M
parameters, much larger than DAVE-2's CNN.
Using such large CNN models will be challenging on
resource constrainted embedded computing platforms, especially for
real-time applications such as self-driving cars.

@@

While continuing performance improvements in embedded computing
platforms will certainly make processing these complex CNNs faster,
another actively investigated approrach is to reduce the required
computational complexity itself.
%% When a DNN is deployed in those implementations with limited
%% resources, such as memory and power, the floating-point operations
%% involved in the large matrix multiplications are a burdensome task.
Many recent advances in network compression have shown promising results
in reducing such computational cost during the feedforward
process. The fundamental assumption in those techniques is that the
DNNs are redundant in their structure and representation. For example,
network pruning can thin out the network and provides a more condensed
topology~\cite{han2015deep}.

Another common compression method is to reduce the
quantization level of the network parameters, so that arithmetic
defined with the floating-point operations are replaced with low-bit
fixed-point counterparts. To this end, single bit quantization of the
network parameters or ternary quantization have been recently proposed
~\cite{hwang2014fixed,soudry2014expectation,kim2016bitwise,rastegari2016xnor,hubara2016binarized,beauchamp2006embedded,govindu2004analysis}.
In those networks, the inner product between the
originally real-valued parameter vectors is defined with XNOR followed
by bit counting, so that the network can greatly minimize the
computational cost in the hardware implementations. This drastic
quantization can produce some additional performance loss, but those
new binarized or ternarized systems provide a simple quantization
noise injection mechanism during training so that the additional error
is minimized to an acceptable level.

The XNOR operation and bit counting have been known to be efficient in
hardware implementations. In~\cite{rastegari2016xnor}, it was shown
that the binarized convolution can substitute the expensive
convolutional feedforward operations in a regular Convolutional Neural
Network (CNN), %% by using only about 1.5\% of the memory space, while
providing 20 to 60 times faster feedforward. Binary weights were also
able to provide 7 times faster feedforward than a floating-point
network for the hand written digit recognition task as well as 23
times faster matrix multiplication tasks on
GPU~\cite{hubara2016binarized}.
Moreover, FPGA implementations showed that the XNOR operation is 200
times cheaper than floating-point multiplications with 
single precision~\cite{beauchamp2006embedded,govindu2004analysis}.
XNOR-POP is another hardware implementation that reduced
the energy consumption of a CNN by 98.7\%~\cite{jiang2017xnor}.

These research efforts are expected to make complex DNNs more
accessible for a wider range of real-time embedded systems. We plan to
investigate the feasibility of these approaches in the context of
DeepPicar so that we can use even more resource constrained
micro-controller class computing platforms.

>>>>

There are several relatively inexpensive RC-car based autonomous car
testbeds. MIT's RaceCar~\cite{shin2017project} and UPenn's
F1$/$10~\cite{upennf1tenth} are both based on a Traxxas 1/10 scale RC
car and a NVIDIA Jetson multicore computing platform, which
is equipped with many sophisticated sensor packages, such as a lidar.
However, they both cost more than \$3,000, requiring a considerable
investment. DonkeyCar~\cite{donkeycar} is similar to our DeepPicar as
it also uses a Raspberry Pi 3 and a similar CNN for end-to-end
control, although it costs more (about \$200).
%% because it uses a more
%% expensive 1/16 scale RC car than our 1/24 scale one.
The main contribution of our paper is in the detailed analysis of
computational aspects of executing a CNN-based real-time control
workload on diverse embedded computing platforms.

In this paper, we have analyzed real-time performance of a real-world
CNN, which was used in NVIDIA's DAVE-2 self-driving
car~\cite{Bojarski2016}, on a low-cost raspberry pi 3 quad-core 
platform and other embedded multicore platforms. It should be noted, 
however, that DAVE-2's CNN is relatively small compared
to recent state-of-the-art CNNs, which are increasingly larger and
deeper. For example, the CNN based object detector models evaluated in
Google's recent study~\cite{huang2017speed} have between 3M to 54M
parameters, and are much larger than DAVE-2's CNN.
Using such large CNN models will be challenging on
resource constrainted embedded computing platforms, especially for
real-time applications such as self-driving cars.

@@

While continuing performance improvements in embedded computing
platforms will certainly make processing these complex CNNs faster,
another actively investigated approach is to reduce the required
computational complexity itself.
%% When a DNN is deployed in those implementations with limited
%% resources, such as memory and power, the floating-point operations
%% involved in the large matrix multiplications are a burdensome task.
Many recent advances in network compression have shown promising results
in reducing such computational costs during the feedforward
process. The fundamental assumption in those techniques is that the
CNNs are redundant in their structure and representation. For example,
network pruning can thin out the network and provides a more condensed
topology~\cite{han2015deep}.

Another common compression method is to reduce the
quantization level of the network parameters, so that arithmetic
defined with floating-point operations are replaced with low-bit
fixed-point counterparts. To this end, single bit quantization of the
network parameters or ternary quantization have been recently proposed
~\cite{hwang2014fixed,soudry2014expectation,kim2016bitwise,rastegari2016xnor,hubara2016binarized,beauchamp2006embedded,govindu2004analysis}.
In those networks, the inner product between the
originally real-valued parameter vectors is defined with XNOR followed
by bit counting, so that the network can greatly minimize the
computational cost in the hardware implementations. This drastic
quantization can produce some additional performance loss, but those
new binarized or ternarized systems provide a simple quantization
noise injection mechanism during training so that the additional error
is minimized to an acceptable level.

The XNOR operation and bit counting have been known to be efficient in
hardware implementations. In~\cite{rastegari2016xnor}, it was shown
that the binarized convolution could substitute the expensive
convolutional feedforward operations in a regular Convolutional Neural
Network (CNN), %% by using only about 1.5\% of the memory space, while
providing 20 to 60 times faster feedforward. Binary weights were also
able to provide 7 times faster feedforward than a floating-point
network for the hand written digit recognition task as well as 23
times faster matrix multiplication tasks on a
GPU~\cite{hubara2016binarized}.
Moreover, FPGA implementations showed that the XNOR operation is 200
times cheaper than floating-point multiplications with 
single precision~\cite{beauchamp2006embedded,govindu2004analysis}.
XNOR-POP is another hardware implementation that reduced
the energy consumption of a CNN by 98.7\%~\cite{jiang2017xnor}.

These research efforts are expected to make complex CNNs more
accessible for a wider range of real-time embedded systems. We plan to
investigate the feasibility of these approaches in the context of
DeepPicar so that we can use even more resource constrained
micro-controller class computing platforms.
